Containers and R: the Rockerverse and beyond

### 2019-09-18 manuscript.Rmd
restructure after telco 2

# Containerization and Rocker

@@

# R Packages

## Packages for automation and packaging [@nuest]

- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockertest`](https://github.com/traitecoevo/dockertest/)

### liftr

@@

- [`rize` for Shiny](https://github.com/cole-brokamp/rize)
- [`dockerfiler`](https://github.com/ColinFay/dockerfiler)

## Packages for control and provisioning [@nuest]

- [`stevedore`](https://github.com/richfitz/stevedore)
- [`babelwhale`](https://cran.r-project.org/web/packages/babelwhale/index.html): Running a Docker from R with Singularity or Docker as back-end. This is really useful in HPC environments where a user might not have root access but is able to install Singularity instead of Docker. [@rcannood]
- [`docker`](https://bhaskarvk.github.io/docker/)
- `RSelenium`
- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `docker_run`)
  - I don't think `docker_run` is main focus for containers for this package, its more how it uses Docker to launch custom environments that can be templated (e.g. rstudio, shiny etc.) and for parrallisation e.g. `library(future)`  using GCE VMs as endpoints. [@MarkEdmondson1234]
- [`analogsea`](https://github.com/sckott/analogsea)
- [`harbor`](https://github.com/wch/harbor/)
- [`dockermachine`](https://github.com/cboettig/dockermachine)

@@

@@

- Missing pieces?
- consolidation (e.g. via packages using `dockerfiler` and `stevedore`)
- Common themes
  - reproducibility
- will knowledge about containers continue to spread?
- what is needed for even more containers with R?
- the ability to move processing between services easily (e.g locally, one cloud providers VM, another cloud provider's Container-as-a-Service)
- ...

>>>>

# Containerization and Rocker [NN]

@@

# R Packages
<!-- from packages closer to Docker to second level packages to more complex ones -->

## Clients
<!-- first order packages, talking directyl to Docker -->

- [`stevedore`](https://github.com/richfitz/stevedore)
- [`docker`](https://bhaskarvk.github.io/docker/)
- [`harbor`](https://github.com/wch/harbor/)
- [`dockermachine`](https://github.com/cboettig/dockermachine)

## Capture and create environments [@nuest]
<!-- second order packages -->

- [`dockerfiler`](https://github.com/ColinFay/dockerfiler/)
- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockertest`](https://github.com/traitecoevo/dockertest/)
- [`rize` for Shiny](https://github.com/cole-brokamp/rize/)

@@

@@

## Deployment to cloud services [@nuest, @richfitz, @rcannood]

- [`babelwhale`](https://cran.r-project.org/web/packages/babelwhale/index.html): Running a Docker from R with Singularity or Docker as back-end. This is really useful in HPC environments where a user might not have root access but is able to install Singularity instead of Docker. [@rcannood]
- `RSelenium`
- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `docker_run`)
  - I don't think `docker_run` is main focus for containers for this package, its more how it uses Docker to launch custom environments that can be templated (e.g. rstudio, shiny etc.) and for parrallisation e.g. `library(future)`  using GCE VMs as endpoints. [@MarkEdmondson1234]
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)

@@

- next steps
  - consolidation?
- support more complex set-ups with ready-to-use stacks, e.g. GPU, AI, ML
- Missing pieces?
- consolidation (e.g. via packages using `dockerfiler` and `stevedore`)
- Common themes
  - reproducibility
- will knowledge about containers continue to spread?
- what is needed for even more containers with R?
- the ability to move processing between services easily (e.g locally, one cloud providers VM, another cloud provider's Container-as-a-Service)
- running RStudio in a container is a common approach for reproducibility - _is there a need to run/control containers from a UI integrated in RStudio?_
- moving away from Docker focus towards containers?
  - is there a need for a wrapper for [podman's Varlink-API](https://github.com/containers/libpod/blob/master/API.md) (similar to stevedore) and buildah to manage containers and images respectively with alternative software stacks?
  - will there be control packages for Singularity support?
- ...
